Pre-trained BERT helps with dialogue act recognition, but not very much

### 2020-12-07 reports/ACL2020 Paper/starsem_reviews-response-notes.md

General comments

# Review 1.
- Although we are only concerned with text in this work, we agree that text is not enough to capture meaning of laughter. These experiments investigate how this model deals with dialogue, given that it was pre-trained on text data. Giving the model acoustic information would undermine that goal.
- We intentionally use a very simple model, which is partly why the description is so short. We will make our code available when the work is no longer under anonymity restrictions.
- DAR is not as appropriate as NLU for laughter?? 
- fix:
  - change to the 1st edition of Austin
  - I also wouldn't say that conditional random fields are more
    "sophisticated" than most flavors of RNNs. -- clarify
- reformulate the contributions
- laughter in non-dialogue?
- why single symbol, fix in Data (ref ginzburg/chiara) but we have
  just a single type.
- why laughter is helpful? restate that in the discussion

idea: can laughter help with repair? 
# Review 2.
- Yes -- $s^t$ is the speaker token and $w_n$ is the n'th word token. Thanks for noticing this omission. 
- This is indeed an interesting discrepancy between the effect of laughter on the CNN and BERT models. We don't have a hypothesis at the moment, but we plan to investigate further. Looking at the DA tag-specific performance may provide a clue.
- The opening two paragraphs get off on the wrong foot for me. The
  first sentence is too often repeated in papers like this, and the
  rest of these two paragraphs seems to under-state the extent to
  which there is prior work on using neural models for dialogue.
  - try to re-write first two para, more laughter and multi-modal dialogue
- label outliers on Fig 1
- say that non-verbal is not accounted for in the Task 1
- Fig 3. Plot: P(qw|qh) - P(qw) instead
- numbers for qh|qw improvement
- Fig 2: utterance encoder, context representation, softmax
- Table 6, add F1
- Experiment 2 -> add L/NL condition
- Experiment 3 -> as well add laughters

# Review 3.
- Thank you for your questions suggesting avenues for deeper analysis.
- Please see figure 4 in the appendix for the affect of laughter on performance for particular dialogue acts.
- We discuss linguistic work on the functional role of laughter and its relationship to our results in section 4.1.
- say something about the split. How is it different from the
  "canonical split" (optional)
- CNN: "a simple and popular baseline for text classification or
  encoding texts"

# Review 4.
- Thank you for these constructive comments.

### 2021-03-10 reports/IWCS2021/iwcs-short-paper.tex
iwcs draft

%
% File acl2020.tex
%
%% Based on the style files for ACL 2020, which were
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2020}
\usepackage{times}
\usepackage{latexsym}
\renewcommand{\UrlFont}{\ttfamily\small}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{natbib}
% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

\usepackage{tabularx}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}

@@

\title{Can BERT learn dialogue-specific phenomena? The case of laughter in dialogue act recognition. }

@@

\begin{document}
\maketitle
\begin{abstract}
   In this work we investigate whether transformer models can learn to represent phenomena specific to dialogue.
   Laughter tokens, which are not present in pre-trained BERT models, can be learned with additional pre-training and fine-tuning on dialogue data. 
   We use dialogue act recognition as a proxy task to demonstrate that laughter can help to disambiguate between certain communicative intents.
   We find that pre-training and fine-tuning procedures are useful for learning laughter representations, and explore how well the model adapts to the dialogue domain.
\end{abstract}

% INTRODUCTION

@@

One important dialogical feature missing from text data is laughter, which is ubiquitous in our everyday interactions.
We observe that in the Switchboard dialogue corpus laughter comes about every 200 tokens.
Laughter also relates to the discourse structure of dialogue and can relate to a \emph{laughable}, which can be an event or an entity in the discourse (or external perceived event as well).
Laughter can precede, follow or overlap the laughable, with the time alignment between the laughter and laughable dependent on who produces the laughable and the form of the laughter \citep{tian2016we}.
Laughter can act as a contextual feature for determining the sincerity of an utterance, e.g. when detecting sarcasm \citep{tepperman2006yeah}.

@@

We use dialogue act recognition (DAR) as a proxy task, since both the internal content of contributions, and their sequential structure have bearing on this task. DAR is the task of assigning a dialogue act tag to each utterance in the dialogue, according to dialogue act it performs.
Using this task we investigate whether transfer learning techniques based on pre-trained language models, such as BERT \citep{devlinBERTPretrainingDeep2018}, are useful for dialogue. 
By looking at laughter, which is certainly missing in the stadnard pre-training corpus, we can determine whether the pre-training is used in the representation of dialogue-specific features.

\paragraph{Contributions}
The main results of this paper have to do with to the usefulness of language model fine-tuning for DAR, the impact of laughter on DAR, and the relationship between the two.
\begin{itemize}
  \item Laughter is useful for dialogue act recognition, and its impact varies across different dialogue acts (\S\ref{sec:experiment1}).
  \item During fine-tuning, BERT learns to represent laughter, a dialogical feature not seen in pre-training (\S\ref{sec:experiment1} and \S\ref{sec:experiment2}).
  \item Standard BERT pre-training is useful for DAR, but the model performs poorly without fine-tuning (\S\ref{sec:experiment2})
  \item Further pre-training with data from the target domain shows promise for dialogue, but we find mixed results when pre-training with a larger corpus of dialogical data from outside the target domain (\S\ref{sec:experiment3}).
    % TODO: we did that but what can we conclude? Perhaps that the nature of the data is very different, so it is not very helpful..?
  \end{itemize}

@@

\paragraph{Preprocessing}
% - Preprocessing: remove disfluencies, acronyms and speaker tokens in AMI, removing laughter % Bill

@@

Utterances are tokenized using with a word piece tokenizer \citep{Wu2016} with a vocabulary of 30,000.
We add a special laughter token to the vocabulary and map all transcribed laughter to that token.
We also add five speaker tokens and prepend each utterance with a speaker token that uniquely identifies the corresponding speaker within that dialogue.

\paragraph{Laughs in the data}
The distribution of laughs in different dialogue acts has a rather uniform shape with a few outliers (Figure~\ref{fig:box}). 
Figure~\ref{fig:by-da} in Appendix~\ref{sec:suppl} provides a breakdown of dialogue act depending on whether current or adjacent utterances (the preceding utterance and the following one are counted as adjacent) contain laughter.\footnote{Note that SWDA has a ``Nonverbal'' dialogue act which is misleading with respect to laughter, because utterances only containing a single laughter token fall into this category, however they can serve, for example, to acknowledge a statement or to deflect a question  \citep{mazzocconi2019phd}.}

\begin{figure}
\centering
\begin{subfigure}[b]{0.55\textwidth}
  \includegraphics[width=0.9\linewidth]{img/box-swda.pdf}
  \caption{SWDA}
  \label{fig:box-swda}
\end{subfigure}

\begin{subfigure}[b]{0.55\textwidth}
  \includegraphics[width=0.9\linewidth]{img/box-ami.pdf}
  \caption{AMI}
  \label{fig:box-ami}
\end{subfigure}

\caption{Box plots for dialogue acts associated with laughter. On the left: proportion of DAs having laughter in one of the adjacent utterances, on the right: proportion of DAs containing laughter. }
\label{fig:box}
\end{figure}

@@

We also experiment with a much larger a corpus constructed from OpenSubtitles \citep{Lison2016}.
We used a manually constructed list of words frequenty used to refer to laughter in subtitles and replaced every occurance of one of these words with the special laughter token. 
Then, we collected every English-language subtitle file in which at least 1\% of the utterances contained laughter (about 11\% of the total).
Because utterances are not labeled with speaker in the OpenSubtitles corpus, we randomly assigned a speaker token to each utterance to maintain the format of the other dialouge corpora.

@@

\begin{table}[ht]
\centering
\begin{tabular}{@{}llll@{}}
\toprule
       & tokens & \% laughs & epochs\\ \midrule
AMI           & 1M     & 1.3             & 10     \\
SWBD          & 1.5M   & 0.5             & 10     \\
OpenSubtitles & 350M   & 0.3             & 1      \\ \bottomrule
\end{tabular}
\caption{Pre-training corpora, showing laughter representation as a percentage of total tokens and number of pre-training epochs generated.}\label{tab:pretraining-corpora}
\end{table}

@@

\section{Experiments}
\subsection{Experiment 1: Impact of laughter} \label{sec:experiment1}   % Vlad
In the first experiment we investigated whether laughter, as an example of a dialogue-specific signal, is a helpful feature for DAR.
Therefore, we train another version of each model: one that containing laughs (\texttt{L}) and one with laughs left out (\texttt{NL}), and compare their performances in DAR task.
Table~\ref{table:laughter-total-acc} compares the results from applying the models with two different utterance encoders (\texttt{BERT}, \texttt{CNN}).

First of all, we can see that the \texttt{BERT} utterance encoder outperforms \texttt{CNN} in all the cases.
From observing the effect of laughters, we can see differences in performance depending on the dialogue act, regardless how often laughter occurs in the current or adjacent utterances (see Figure~\ref{fig:by-da} in Appendix~\ref{sec:suppl}).
The strongest evidence for laughter as a helpful feature was found in SWDA for the \texttt{BERT} utterance encoder, where macro-F1 score increases by 7.89 percentage points.

\begin{table}
  \centering
  \begin{tabular}{@{}lcccc@{}}
    \toprule
                      & \multicolumn{2}{c}{SWDA} & \multicolumn{2}{c}{AMI-DA} \\ \midrule
                      & F1    & acc.  & F1    & acc.       \\ 
    \texttt{BERT-NL}  & 38.10 & 77.07 & 49.09 & 67.06       \\ 
    \texttt{BERT-L}   & 45.99 & 76.93 & 50.17 & 67.12       \\ \midrule
    \texttt{CNN-NL}   & 37.23 & 75.08 & 38.37 & 63.46        \\
    \texttt{CNN-L}    & 27.59 & 75.40 & 37.94 & 64.30        \\ \midrule
    Majority class    & 0.78  & 33.56 &  1.88 & 28.27      \\ \bottomrule

  \end{tabular}
  \caption{Comparison of macro-average F1 and accuracy depending on using laughter on the training phase. }
  \label{table:laughter-total-acc}
\end{table}

Confusion matrices (Figure~\ref{fig:swda-cm}) provide some food for thought. Most of the misclassifications fall into the majority classes, such as \emph{sd} (Statement-non-opinion), in the left edge of the matrix. However, there are some important exceptions, such as \emph{rhetorical questions}, that are misclassified as other forms of questions due to their surface question-like form.
Importantly, laughter helps to classify rhetorical questions correctly, this is because in a conversation it can be used as a device to cancel seriousness or sincerity \citep{ginzburg2015understanding,tepperman2006yeah}, in this case, of a question.
Therefore, questions, like the one we show in the example in Table~\ref{table:example-qh}, are easier to disambiguate with the help of laughter.
\begin{table}
      \small
  \centering
  \begin{tabularx}{\linewidth}{llL}
    \toprule
    Speaker & DA & Utterance \\ \midrule
        \multicolumn{3}{c}{\emph{(Talking about hobbies)}}\\
    B   & sd	&  Um, as far as spare time, they talked about, \\
    B	& \% & I don't, + I think, \\
    B	& qh & who has any spare time \texttt{<laughter>}? \\
    A	& x & \texttt{<laughter>}.\\
             \bottomrule
  \end{tabularx}
  \caption{Example from the SWDA corpus (sw3735). B's contribution \emph{qh} (Rhetorical question) is misinterpreted as \emph{qw} (Wh-question) by the BERT model without laughs in training data. }
  \label{table:example-qh}
\end{table}

  \begin{figure}
  \centering
  \includegraphics[width=1.0\linewidth]{img/swda-cm.pdf}
  \caption{Confusion matrices for \texttt{BERT-NL} (top) vs \texttt{BERT-L} (bottom); SWDA corpus. Solid line shows the improvement in classification of rhetorical questions.}
    \label{fig:swda-cm}
  \end{figure}

In order to further investigate the differences between the models and datasets, we looked at the dialogue act groups.
Mapping ground-truth labels and predictions into four groups allows direct comparison between model performances in SWDA and AMI-DA.
Table~\ref{table:laughter-group-acc} compares accuracies between the two main groups: Forward-Communicative-Function and Backward-Communicative-Function (82.3\% of all DAs in SWDA and 76.7\% in AMI-DA).
Our models tend to perform better for the Forward-Communicative-Function group for both corpora irrespective of the presence of laughs.
Also, our models perform better on the SWDA corpus.
Overall performance in these groups appears to be unaffected by presence of laughs.

\begin{table}
  \centering
  \begin{tabular}{@{}llcc@{}}
    \toprule
    &                  & Forw.-CF  & Backw.-CF   \\ \midrule
    \multirow{4}{*}{SWDA}    &\texttt{BERT-NL}  & 92.54 & 91.67       \\ 
                             &\texttt{BERT-L}   & 95.42 & 90.91       \\ 
                             &\texttt{CNN-NL}   & 93.30 & 90.85        \\
                             &\texttt{CNN-L}    & 94.64 & 90.13        \\ \midrule
    \multirow{4}{*}{AMI-DA}  &\texttt{BERT-NL}  & 87.77 & 80.83       \\ 
                             &\texttt{BERT-L}   & 87.36 & 79.87       \\
                             &\texttt{CNN-NL}   & 91.85 & 71.21        \\
                             &\texttt{CNN-L}    & 91.25 & 71.28        \\    \bottomrule
  \end{tabular}
  \caption{Comparison of the accuracies among the two main dialogues act groups, depending on the utterance model and the presence of laughter.}
  \label{table:laughter-group-acc}
\end{table}

@@

\paragraph{Fine-tuning and laughter}
Since the laughter token doesn't appear in the original BERT vocabulary, only fine-tuned models can meaningfully make use of laughter in utterance representations. 
This gives us another opportunity to assess whether laughter is useful in dialogue act recognition.
We find that although 4.6\% of utterances in SWDA contain laughter, 
7.3\% of utterances misclassified by the frozen model but correctly classified by the fine-tuned model contain laughter, 
suggesting that the fine-tuned encoder makes use of laughter. 
For AMI-DA, the effect is less pronounced, but still present: 
overall, 8.5\% of utterances contain laughter, 
compared to 9.6\% of utterances where fine-tuning makes a difference.

@@

\section{Discussion} % add items to discuss
On the question of laughter impact, this study found that laughter is more helpful in SWDA corpus than in AMI-DA.
Due to the nature of interactions over the phone, SWDA dialogue participants can not rely on visual signals, such as gestures and facial expressions.
Our results support the hypothesis that in SWDA, vocalizations such as laughter are more pronounced and therefore more helpful in disambiguating dialogue acts.

@@

There is abundant room for further progress in determining how other speech-related information, such as prosody and disfluencies, can be incorporated into the DAR model.
\citet{stolckeDialogueActModeling2000} showed that dialogue acts can have specific prosodic manifestations that can be used to improve dialogue act classification.
Furthermore, in relation to laughter, its form (duration, arousal level) can be informative about its function and position w.r.t. to laughable \citep{tian2016we,mazzocconi2019phd}.
Incorporating such information is crucial if models pre-trained on large-scale text corpora are to be adapted for use in dialogue applications.

@@

%\begin{itemize}
  %\item RoBERTa training scheme finds that next sentence prediction is of little importance. Is this true for us too? (suspect not due to discourse importance). see: https://github.com/huggingface/transformers/issues/1622
  %\item Issues related to working with transcribed speech: Different kinds of laughter are all represented by the same token. Does transcription orthography capture prosody? We know prosidy is useful for DAR (Jurafsky).
  %\item future work: pre-training on very large un-labeled dialogue corpora. How does text chat transfer to the spoken/transcribed setting?
  %\item Laughter is more helpful in SWDA than in AMI-DA. 
  %\item The point above can also explain why our models better on SWDA: just a larger proportion of stuff that interlocutors rely on is in the transcription.
%\end{itemize}

\bibliography{acl2020}
\bibliographystyle{acl_natbib}
\appendix
\onecolumn
\section{Supplementary materials}\label{sec:suppl}

\begin{figure*}[h!]
  \centering
  \includegraphics[width=\linewidth]{img/SWDA-bertLvsNL.pdf}
  \includegraphics[width=\linewidth]{img/AMI-DA-bertLvsNL.pdf}
  \caption{Change in accuracy for each dialogue act (\texttt{BERT-NL} vs \texttt{BERT-L}). Positive changes when adding laughter (\texttt{BERT-L}) are shown in blue. Vertical bars indicate how often dialogue act is associated with laughter. Top chart: SWDA, Bottom chart: AMI-DA. }
    \label{fig:by-da}
  \end{figure*}

\begin{figure*}[h!]
  \centering
  \includegraphics[width=\linewidth]{img/swda-brf.pdf}
  \includegraphics[width=\linewidth]{img/ami-brf.pdf}
  \caption{F1 scores by dialogue act for BERT with standard pre-training and DAR fine-tuning (\texttt{BERT-FT}) vs.~the same model without pre-training (\texttt{BERT-RI}) and without fine-tuning (\texttt{BERT-FZ}). Dialogue acts are ordered with the most common on the left.}
    \label{fig:f1-by-da}
  \end{figure*}

>>>>

%
% File acl2021.tex
%
%% Based on the style files for EMNLP 2020, which were
%% Based on the style files for ACL 2020, which were
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2021}
\usepackage{times}
\usepackage{latexsym}
\renewcommand{\UrlFont}{\ttfamily\small}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

\usepackage{booktabs}
\usepackage{tabularx}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}

@@

\title{Pre-trained BERT helps with dialogue act recognition, but not very much}

@@

\begin{document}
\maketitle

\begin{abstract}
This document contains the instructions for preparing a manuscript for the proceedings of ACL-IJCNLP 2021.
The document itself conforms to its own specifications, and is therefore an example of what your manuscript should look like.
These instructions should be used for both papers submitted for review and for final versions of accepted papers.
Authors are asked to conform to all the directions reported in this document.
\end{abstract}

%INTRODUCTION

@@

@@

We use dialogue act recognition (DAR) as a proxy task, since both the internal content of contributions, and their sequential structure have bearing on this task. DAR is the task of assigning a dialogue act tag to each utterance in the dialogue, according to dialogue act it performs.
Using this task we investigate whether transfer learning techniques based on pre-trained language models, such as BERT \citep{devlinBERTPretrainingDeep2018}, are useful for dialogue.

\paragraph{Contributions}
The main results of this paper have to do with to the usefulness of language model fine-tuning for DAR.
\begin{itemize}
  \item Standard BERT pre-training is useful for DAR, but the model performs poorly without fine-tuning (\S\ref{sec:experiment2})
  \item Further pre-training with data from the target domain shows promise for dialogue, but we find mixed results when pre-training with a larger corpus of dialogical data from outside the target domain (\S\ref{sec:experiment3}).
  \end{itemize}

@@

\paragraph{Preprocessing}
% - Preprocessing: remove disfluencies, acronyms and speaker tokens in AMI, % Bill

@@

Utterances are tokenized using with a word piece tokenizer \citep{Wu2016} with a vocabulary of 30,000.
We also add five speaker tokens and prepend each utterance with a speaker token that uniquely identifies the corresponding speaker within that dialogue.

@@

We also experiment with a much larger a corpus constructed from OpenSubtitles \citep{Lison2016}.
Because utterances are not labeled with speaker in the OpenSubtitles corpus, we randomly assigned a speaker token to each utterance to maintain the format of the other dialouge corpora.

@@

@@

@@

@@

\section{Discussion} % add items to discuss

@@

There is abundant room for further progress in determining how other speech-related information, such as prosody and disfluencies, can be incorporated into the DAR model.
\citet{stolckeDialogueActModeling2000} showed that dialogue acts can have specific prosodic manifestations that can be used to improve dialogue act classification.
Incorporating such information is crucial if models pre-trained on large-scale text corpora are to be adapted for use in dialogue applications.

@@

\bibliography{acl2020}
\bibliographystyle{acl_natbib}
\appendix
\onecolumn
\bibliographystyle{acl_natbib}
\bibliography{iwcs-short-paper}

%\appendix
